---
title: "508 Midterm"
author: "Caesar"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidycensus)
library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots
```

```{r}
data <- st_read("https://raw.githubusercontent.com/mafichman/musa_5080_2023/main/Midterm/data/2023/studentData.geojson")
```

```{r}
variables <- c('objectid',
               'census_tract',
               'central_air',
               'depth',
               'exempt_land',
               'exterior_condition',
               'fireplaces',
               'frontage',
               'fuel',
               'garage_spaces',
               'garage_type',
               'general_construction',
               'interior_condition',
               'number_of_bathrooms',
               'number_of_bedrooms',
               'number_of_rooms',
               'number_stories',
               'off_street_open',
               'parcel_shape',
               'quality_grade',
               'sale_price')
```

```{r}
f_df <- data[, variables]
```

```{r}
ex_his <- ggplot(f_df, aes(x = exterior_condition)) +
  geom_histogram(binwidth = 1, fill = "grey", color = "black") +
  labs(title = "exterior", x = "Level", y = "Frequency")

in_his <- ggplot(f_df, aes(x = interior_condition)) +
  geom_histogram(binwidth = 1, fill = "grey", color = "black") +
  labs(title = "interior", x = "Level", y = "Frequency")

qua_his <- ggplot(f_df, aes(x = quality_grade)) +
  geom_histogram(stat='count',fill = "grey", color = "black") +
  labs(title = "quality", x = "Level", y = "Frequency")

par_his <- ggplot(f_df, aes(x = parcel_shape)) +
  geom_histogram(stat='count',fill = "grey", color = "black") +
  labs(title = "parcel", x = "Level", y = "Frequency")

nos_his <- ggplot(f_df, aes(x = number_stories)) +
  geom_histogram(stat='count',fill = "grey", color = "black") +
  labs(title = "num_story", x = "Level", y = "Frequency")

nobed_his <- ggplot(f_df, aes(x = number_of_bedrooms)) +
  geom_histogram(stat='count',fill = "grey", color = "black") +
  labs(title = "num_bed", x = "Level", y = "Frequency")

nobath_his <- ggplot(f_df, aes(x = number_of_bathrooms)) +
  geom_histogram(stat='count',fill = "grey", color = "black") +
  labs(title = "num_bath", x = "Level", y = "Frequency")

gencon_his <- ggplot(f_df, aes(x = general_construction)) +
  geom_histogram(stat='count',fill = "grey", color = "black") +
  labs(title = "num_construction", x = "Level", y = "Frequency")

grid.arrange(ex_his, in_his, qua_his, par_his, nos_his,nobed_his, nobath_his,gencon_his, ncol = 4)
```

```{r}
sale_box <- ggplot(f_df, aes(y = sale_price)) +
  geom_boxplot() +
  labs(title = "Boxplot of sale price",
       x = "sale price",
       y = "Value")

ostrt_box <- ggplot(f_df, aes(y = off_street_open)) +
  geom_boxplot() +
  labs(title = "Boxplot of off_street_open",
       x = "offstreet_space",
       y = "Value")

dep_box <- ggplot(f_df, aes(y = depth)) +
  geom_boxplot() +
  labs(title = "Boxplot of depth",
       x = "depth",
       y = "Value")

fr_box <- ggplot(f_df, aes(y = frontage)) +
  geom_boxplot() +
  labs(title = "Boxplot of frontage",
       x = "frontage",
       y = "Value")

grid.arrange(sale_box, ostrt_box, dep_box, fr_box, ncol = 4)
```

```{r outlier process}
df_pro <- f_df
df_pro$frontage <- pmin(df_pro$frontage, 500)
df_pro$depth <- pmin(df_pro$depth, 600)
df_pro$sale_price <- pmin(df_pro$sale_price,4000000)
df_pro$off_street_open <- pmin(df_pro$off_street_open,10000)
df_pro$number_of_bedrooms <- pmin(df_pro$number_of_bedrooms,10)
df_pro$number_of_bathrooms <- pmin(df_pro$number_of_bathrooms,5)
```

```{r drop column with large missing data}
df_pro <- subset(df_pro,select = -c(number_of_rooms,quality_grade,general_construction,exempt_land,fireplaces))
```

```{r deal with missing data}
df_pro$central_air[is.na(df_pro$central_air)] <- 'N'
df_pro$depth[is.na(df_pro$depth)] <- 0
df_pro <- df_pro %>%
  filter(!is.na(exterior_condition))%>%
  filter(!is.na(exterior_condition))%>%
  filter(!is.na(interior_condition))%>%
  filter(!is.na(number_of_bedrooms))%>%
  filter(!is.na(number_stories))%>%
  filter(!is.na(number_of_bathrooms))%>%
  filter(!is.na(off_street_open))%>%
  filter(!is.na(garage_spaces))
```

```{r}
acs_variable_list.2021 <- load_variables(2021, #year
                                         "acs1", #five year ACS estimates
                                         cache = TRUE)
acs_variable_list.2020 <- load_variables(2020, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)
```

```{r results='hide'}
tracts21 <-  
  get_acs(geography = "tract",
          variables = c("B01001_001E", # ACS total Pop estimate
              'B02001_002E', # Estimate!!Total:!!White alone
              'B02001_003E', # Estimate!!Total:!!Black or African American alone
              'B02001_005E', # Estimate!!Total:!!Asian alone
              'B19013A_001E',# Median household income White alone
              'B19013B_001E',# Median household income Black alone
              'B19013D_001E',# Median household income Asian alone
              'B19013_001E', # Median household income 
              'B01002_001E', # Median Age by Sex 
              "B25002_001E", # Estimate of total housing units
              "B25002_003E", # Number of vacant housing units
              "B06009_006E", # Total graduate or professional degree
              "B15001_050E",
              "B15001_009E",
              "B25058_001E",
              "B06012_002E",
              'B25105_001E', # Median monthly housing costs 
              'B25037_001E', # Median year structure built --!!Total
              'B10058_002E', # Total:!!In labor force
              'B10058_007E', # Not in labor force
              'B08141_017E', # Public transportation:!!No vehicle available
              'B08141_018E', # Public transportation:!!1 vehicle available 
              'B09002_002E', # OWN CHILDREN UNDER 18 YEARS married family
              'B09002_015E', # Female householder
              'B14001_004E'), # Enrolled in kindergarten 
          year=2021, state=42,
          county=101, geometry=TRUE, output = "wide") %>% 
  st_transform('ESRI:102728') %>%

  rename(TotalPop = B01001_001E, 
         White = B02001_002E,
         Black = B02001_003E,
         Asian = B02001_005E,
         MedHHInc_White = B19013A_001E,
         MedHHInc_Black = B19013B_001E,
         MedHHInc_Asian = B19013D_001E,
         MedHHInc = B19013_001E,
         Median_Age_bysex = B01002_001E,
         Total_Housing_Units = B25002_001E,
         Vacant_Housing_Units = B25002_003E,
         Total_Graduate_Prof_Degree = B06009_006E,
         FemaleBachelors = B15001_050E, 
         MaleBachelors = B15001_009E,
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E,
         Med_mon_hscst = B25105_001E,
         Med_built_year = B25037_001E,
         Lab_force = B10058_002E,
         no_lav_force = B10058_007E,
         no_pub_trans = B08141_017E,
         one_pub_trans = B08141_018E,
         own_child = B09002_002E,
         own_child_fe = B09002_015E,
         enroll_kinder = B14001_004E)
```

```{r}
df_pro <- df_pro %>%
  st_transform('ESRI:102286')
tracts21 <- tracts21 %>%
  st_transform('ESRI:102286')
```

```{r drop M column}
tracts21 <- tracts21[, !grepl("M$", names(tracts21))]
```

```{r tracts data clean}
mean_in_w <- mean(tracts21$MedHHInc_White,na.rm = TRUE)
tracts21$MedHHInc_White[is.na(tracts21$MedHHInc_White)] <- mean_in_w

mean_in_b <- mean(tracts21$MedHHInc_Black,na.rm = TRUE)
tracts21$MedHHInc_Black[is.na(tracts21$MedHHInc_Black)] <- mean_in_b

mean_in_a <- mean(tracts21$MedHHInc_Asian,na.rm = TRUE)
tracts21$MedHHInc_Asian[is.na(tracts21$MedHHInc_Asian)] <- mean_in_a

mean_in <- mean(tracts21$MedHHInc,na.rm = TRUE)
tracts21$MedHHInc[is.na(tracts21$MedHHInc)] <- mean_in

med_age <- mean(tracts21$Median_Age_bysex,na.rm = TRUE)
tracts21$Median_Age_bysex[is.na(tracts21$Median_Age_bysex)] <- med_age

med_rent <- mean(tracts21$MedRent,na.rm = TRUE)
tracts21$MedRent[is.na(tracts21$MedRent)] <- med_rent

mean_medmonhh <- mean(tracts21$Med_mon_hscst,na.rm = TRUE)
tracts21$Med_mon_hscst[is.na(tracts21$Med_mon_hscst)] <- mean_medmonhh

med_b_yr <- median(tracts21$Med_built_year,na.rm = TRUE)
tracts21$Med_built_year[is.na(tracts21$Med_built_year)] <- med_b_yr
```

```{r}
hh_tract <- st_join(df_pro,tracts21, join = st_within)
```

# The strucuture of housing price modeling
## data prepare 
housing price, census, school, park, crime (the dataset which may have related to the housing price)

## data wrangling 
clean the housing price data (check the variable and keep useful one)
make the geo data (school park crime) into useful variable can be used in prediction
  make pre visualization about the variable created (check regularity)
combine the data into variable of price predict dataset

## principle component analysis 
check the distribution of data to fix the outlier
make correlation test to compare set of variables (like num of knn)
check collinear of variables

## data modeling
use selected varialbes to build inital model 
add spatial lag
shit!

```{r}
crime_path <- 'https://phl.carto.com/api/v2/sql?filename=incidents_part1_part2&format=csv&q=SELECT%20*%20,%20ST_Y(the_geom)%20AS%20lat,%20ST_X(the_geom)%20AS%20lng%20FROM%20incidents_part1_part2%20WHERE%20dispatch_date_time%20%3E=%20%272021-01-01%27%20AND%20dispatch_date_time%20%3C%20%272022-01-01%27'

crime <- read.csv(crime_path)
```

```{r}
crime <- crime%>%
  filter(lat > 0)
crime_df <- st_as_sf(crime, coords = c("lng", "lat"), crs = 4326)
```

```{r}
# Parks & Recreation Program Sites
park <- 'https://opendata.arcgis.com/api/v3/datasets/9eb26a787a6e448ba426eea7f9f0d93a_0/downloads/data?format=geojson&spatialRefId=4326'
park <- st_read(park)
```

```{r}
park2 <- 'https://opendata.arcgis.com/api/v3/datasets/38f251496c10432ca0560e7989fcc316_0/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1'
park2 <- st_read(park2)
```

```{r}
school <- 'https://opendata.arcgis.com/datasets/d46a7e59e2c246c891fbee778759717e_0.geojson'
school <- st_read(school)
```